\documentclass[12pt]{article}

% Packages for basic formatting
\usepackage{amsmath}  % for mathematical symbols
\usepackage{amsfonts} % for fonts
\usepackage{amssymb}  % for more symbols
\usepackage{geometry} % to customize page layout
\usepackage{fancyhdr} % for custom headers
\usepackage{hyperref} % for clickable references
\usepackage{tikz}     % optional, for drawings
\usepackage{graphicx} % optional, for images

\usepackage{ocgx}
\usepackage[shortlabels]{enumitem}
\usepackage{tabularx}

\setlist[enumerate]{font=\bfseries}

% Page dimensions and margins
\geometry{top=1in, bottom=1in, left=1in, right=1in}

% Title info
\title{\textbf{Python Journey} \\ \large Introductory Problem Set}
\date{\today}

\begin{document}

% Title
\maketitle

% Instructions (if any)
\noindent \textbf{Disclaimer} \\\\
In this handout we assume that students:
\begin{itemize}
    \item Have already been introduced to Python functions, conditionals, 
      for and while loops.
    \item Program through scripts instead of using the console.
    \item The objective of these exercises is to reach a simple and intuitive
      solution, which not always corresponds to the most efficient or complete.
\end{itemize}

\subsection*{Exercise 1 - Even Odd} 
\begin{enumerate}[a)]
  \item Develop a function that receives an integer and calculates whether it 
    is odd or even.
  \item Develop a test function for the Even Odd problem.
\end{enumerate}

\subsection*{Exercise 2 - Grade Calculator} 

The following grading system is commonly used in the United States of America:
\begin{table}[!ht]
  \begin{center}
      \begin{tabular}{ | m{7em} | m{7em} | }
          \hline
          Letter Grade  & Number Grade    \\
          \hline
          A             & 90-100          \\
          \hline
          B             & 80-89           \\
          \hline
          C             & 70-79           \\
          \hline
          D             & 60-69           \\
          \hline
          F             & 0-59             \\
          \hline
      \end{tabular}
  \end{center}
\end{table}
\vspace{-1cm}
\begin{enumerate}[a)]
  \item Develop a function that receives an integer and calculates the 
    corresponding letter grade. \textbf{Remember} to verify if the number 
    received is within the desired scale.
  \item Develop a test function for the Grade Calculator problem.
\end{enumerate}

\subsection*{Exercise 3 - List Summation}

Let $a = a_{0}, ..., a_{n-1}$, then $a$ is sequence with $n$ elements, starting from
index $0$ to $n-1$. The sum of the elements in sequence $a$, can be expressed as: 

\[ \sum_{i=0}^{n-1} a_{i} = a_{0} + ... + a_{n-1} \]

\vspace{2mm}
\noindent \textbf{Note} that sequences can be stored as lists in python.
\begin{enumerate}[a)]
  \item Develop a function that receives a list of floats and calculates the
    sum of its elements.
  \item Develop a test function for the List Summation problem.
\end{enumerate}

\subsection*{Exercise 4 - Vowel Count}
\begin{enumerate}[a)]
  \item Develop a function that calculates the number of vowels in a string. \\
    \textbf{Warning}: The comparison between strings is \textbf{case-sensitive}.
  \item Develop a test function for the Vowel Count problem.
\end{enumerate}

\subsection*{Exercise 5 - Multiplication Table}
Multiplication tables are often used when teaching students multiplication for
the first time, one example being:

\begin{table}[!ht]
  \begin{center}
      \begin{tabular}{ | m{7em} | }
          \hline
          Tabuada do 2  \\
          \hline
          $2 * 1 = 2$   \\
          \hline
          $2 * 2 = 4$   \\
          \hline
          $2 * 3 = 6$   \\
          \hline
          $2 * 4 = 8$   \\
          \hline
          $2 * 5 = 10$  \\
          \hline
          $2 * 6 = 12$  \\
          \hline
          $2 * 7 = 14$  \\
          \hline
          $2 * 8 = 16$  \\
          \hline
          $2 * 9 = 18$  \\
          \hline
          $2 * 10 = 20$ \\
          \hline
      \end{tabular}
  \end{center}
\end{table}

\begin{enumerate}[a)]
  \item Develop an interactive program that prints the multiplication table
    for numbers selected by the user one at a time. The user should also be
    able to terminate the program, for instance with the word "end". 
    \textbf{Remember} to verify if the input is valid.
\end{enumerate}

\subsection*{Exercise 6 - Palindromes}
A palindrome is a sequence of symbols that can be read forward and backwards
in the same way. However, one may usually find that some liberties are taken,
for instance with punctuation or formatting. For simplicity, we recommend
only considering truly palindromic sequences of characters, i.e. the characters
must exactly match in both ways, with the only exception being the letter cases, 
if the student wishes. \\

\noindent By our definition:
\noindent \begin{itemize}
  \item "Stop pots" is truly palindromic, if one ignores letter cases;
  \item "Never odd or even" is not truly palindromic, since the 5th character 
    from the start is "r" and from the back is a space.
\end{itemize}

\begin{enumerate}[a)]
  \item Develop a function that determines whether a string is truly palindromic.
  \item Develop a test function for the Palindromes problem.
\end{enumerate}

\subsection*{Exercise 7 - Fibonacci}
There are innumerous ways to calculate the $n^{th}$ number in the Fibonacci 
sequence. However, in this exercise students should use the iterative method, 
which means starting from the first and second Fibonacci numbers until one
reaches the $n^{th}$ number, using a loop.
\begin{enumerate}[a)]
  \item Develop a function that calculates the $n^{th}$ Fibonacci number.
  \item Develop a test function for the Fibonacci problem.
\end{enumerate}

\subsection*{Exercise 8 - Maximum Element}
\begin{enumerate}[a)]
  \item Develop a function that calculates the maximum element of a list of 
    floats. \\ \textbf{Suggestion:} the resolve variable can be initialized as 
    $float(inf)$
  \item Develop a test function for the Maximum problem.
  \item Repeat the previous subproblems for the minimum element.
  \item \textbf{Challenge:} Repeat the previous subproblems for the second-highest
    value of a list.
\end{enumerate}

\subsection*{Exercise 9 - Prime Numbers}
A prime number is any integer greater than one and only be divisible by itself 
and one. The first five prime numbers are: $2, 3, 5, 7, 11$.

\begin{enumerate}[a)]
  \item Develop a function that calculates if a given integer is a prime number.
  \item Develop a test function for the Prime Numbers problem.
\end{enumerate}

\subsection*{Exercise 10 - Inversion}
\begin{enumerate}[a)]
  \item Develop a function that inverts the order of the elements in a list. 
    \textbf{Suggestion:} The result should be a new list.
  \item Develop a test function for the Inversion problem.
\end{enumerate}

\subsection*{Exercise 11 - Remove Duplicates}
\begin{enumerate}[a)]
  \item Develop a function that removes the elements that are duplicated in a 
  list. \textbf{Suggestion:} The result should be a new list.
  \item Develop a test function for the Inversion problem.
\end{enumerate}


\end{document}